---
title: "Multi-Scale Ordination Under Variogram Based Framework"
output: html_document
---

Read in data and select data separated at a specific grain size

```{r}
setwd('C:/Users/Documents/Clay Dustin/Spring 2015/Quantitative Methods/tree_vario/Data')
data = read.csv('./data/cross_comms.csv')
grainSize = 156.25
species.RichnessMatrix = as.matrix(data[which(data$grain==grainSize),-(1:3)])
```

Create the distance matrix between sites
```{r eval=TRUE}
xyCoords = data[,2:3]
dist.Mat = dist(xyCoords)*sqrt(grainSize)
```

Create a list of distance classes 
```{r}
H <- ceiling(dist.Mat/sqrt(grainSize))*grainSize
hmax <- round((max(dist.Mat/2)/grainSize)*grainSize)
H[H>hmax] <- max(H)
dist.Classes = unique(H)
H <- as.matrix(H)
```
Now that we have distinct distance intervals we want to separate the paired sites according to the distance between them.
```{r}
sites.ByDist = list()
for(i in 1:length(dist.Classes))
    sites.ByDist[[i]] = which(H == dist.Classes[i], arr.ind = TRUE)
```

Now that we have a list of unique distance intervals that exist between sites we can generate covariance matrices based on each distance interval.
```{r}
cov.Mat = list()
for(i in 1:length(dist.Classes)){
    diff = species.RichnessMatrix[sites.ByDist[[i]][,1],]-species.RichnessMatrix[sites.ByDist[[i]][,2],]
    cov.Mat[[i]] = matrix(0, nrow = ncol(diff), ncol = ncol(diff))
    for(j in 1:nrow(diff)){
        cov.Mat[[i]] = cov.Mat[[i]] + diff[j,]%*%t(diff[j,])
    }
}
```
Now we construct the total covariance matrix which is the sum of all the covariance matrices by H.
```{r eval=TRUE}
C = Reduce('+',cov.Mat)
```

Next we perform Principal Component Analysis on the summed matrix which results will allow us to identify the main causes for variance in species richness. 
```{r eval=TRUE}
eigen.ForC<-eigen(C)
eigen.values<-eigen.ForC$values
eigen.vectors<-eigen.ForC$vectors
```

Next we extract the proportion of eigenvalue represented at each distance class. This allows us to determine at what H most of the variance reported by the PCA is represented at. 
```{r eval =TRUE}
weighted.Eigenvalue = t(eigen.vectors[1])%*%cov.Mat%*%eigen.vectors[1]
```

#create matrices of complimentarity (where the covariance matrix is not defined for between species variances)
comp.Cov = list()
for(i in 1:length(cov.Mat)){
    comp.Cov[[i]] = diag(diag(cov.Mat[[i]]))
}